---
title: "R for Data Science - Solutions Manual"
author: "Jonathan Mendes de Almeida"
output:
    pdf_document:
      latex_engine: xelatex
---

Solutions to the exercises in __[R for Data Science (Garrett Grolemund & Hadley Wickham)](http://r4ds.had.co.nz/index.html).__

# Chapter 1
No exercises in this chapter.

# Chapter 2
No exercises in this chapter.

# Chapter 3
## Prerequisites

```{r eval = TRUE}
library(tidyverse)
```

### To install tidyverse library:

```{r eval = FALSE}
install.packages('tidyverse')
```

## mpg data frame

```{r eval = TRUE}
mpg = ggplot2::mpg
mpg
```

* **displ**: car's engine size (in litres)
* **hwy**: car's fuel efficiency on the highway (in miles per gallon - mpg).

## 3.2.4 Exercises
> *1.* Run **ggplot(data = mpg)**. What do you see?

```{r eval = TRUE}
ggplot(data = mpg)
```

**R)** An empty plot. To see some nice plots we should add some geom_function to map some points. Add a good caption for each axis is great to make your plot easier to read and understand!

> *2.* How many rows are in **mpg**? How many columns?

```{r eval = TRUE}
nrow(mpg)
ncol(mpg)
```

**R)** 234 rows and 11 columns


Alternative method to check the number of rows and columns of a data frame:

```{r eval = TRUE}
glimpse(mpg)
```


> *3.* What does the __drv__ variable describe? Read the help for ?mpg to find out.
_run ?mpg in RStudio console and check the 'help' tab)_

```{r eval = TRUE}
mpg['drv']
```

**R)** There are 3 possible values for __drv__ variable (**variable** : _description_):
 
* **f**: _front-wheel drive_

* **r**: _rear wheel drive_

* **4**: _4wd_

> *4.* Make a scatterplot of **hwy** vs **cyl**

```{r eval = TRUE}
ggplot(data = mpg, aes(x = hwy, y = cyl)) + geom_point(aes(colour = class))
```

**R)** We add a definition of the data used in x,y axis and add geom_point function mapping these points. Here we are using aes(colour = class) to plot using a different colour for each class present in our data frame (car type: 2seater, compact, midsize, minivan, pickup, subcompact, suv).

> *5.* What happens if you make a scatterplot of *class* __vs__ *drv*? Why is the plot not useful?

```{r eval = TRUE}
ggplot(data = mpg, aes(x = class, y = drv)) + geom_point()
```

**R)** This plot is not useful because **class** and **drv** are factor variables. Each possible value of these two variables is limited by a set (r, f and 4 are the possible values for **drv**  and 2seater, compact, midsize, minivan, pickup, subcompact and suv are the possible values for **class**). This plot is pretty useless to perform a data analysis.









